### 2012-05-23 doc/hmcsa_adjoint_vs_direct/direct_vs_adjoint.tex
Added results figures to adjoint vs direct tech note

\refno{RNSD-01}
\subject{Stochastic Linear Solver Comparison for Monte Carlo Synthetic
Acceleration Method}

@@

%%---------------------------------------------------------------------------%%
\section{Introduction}
The Monte Carlo Synthetic Acceleration (MCSA) method developed by
Evans et. al. \cite{Evans_2009} provides an acceleration to standard
stochastic linear solver methods that are bound by the central limit
theorem, instead giving exponential convergence. To do this, we must
first consider stochastic solutions for linear systems. For the
following linear system

@@

Using this form, we can approximate the terms in the series with a
random walk sequence as outlined in \cite{Hammersley_1964}. There are
two primary ways of achieving this; through a direct sequence and an
adjoint sequence. For the direct sequence, analagous (and
inconveniently so) to an adjoint stochastic transport calculation,
each term in the source is simulated with an equivalent number of
random walk histories. Each step in the random walk sequence will
contribute to the solution in the starting source state. For the
adjoint sequence, analagous to a forward stochastic transport
calculation, the source term is sampled with a set number of random
walk histories, meaning that regions of larger source will be sampled
as a starting state more regularly. In this case, each step in the
random walk sequence will contribute to the state in which it
currently resides, not the current source state. It is this set of
primary differences between the two solution methods that will dictate
solver performance with MCSA.

@@

@@

@@

% \bibliographystyle{notes}
% \bibliography{direct_vs_adjoint}

>>>>

\refno{RNSD-01} \subject{Stochastic Linear Solver Comparison for the
  Monte Carlo Synthetic Acceleration Method}

@@

%%---------------------------------------------------------------------------%%
\section{Introduction}
The Monte Carlo Synthetic Acceleration (MCSA) method developed by
Evans et. al. \cite{evans_2003} provides an acceleration to standard
stochastic linear solver methods that are bound by the central limit
theorem, instead giving exponential convergence. To do this, we must
first consider stochastic solutions for linear systems. For the
following linear system

@@

Using this form, we can approximate the terms in the series with a
random walk sequence as outlined in \cite{hammersley_1964}. There are
two primary ways of achieving this; through a direct sequence and an
adjoint sequence. For the direct sequence, analagous (and
inconveniently so) to an adjoint stochastic transport calculation,
each term in the source is simulated with an equivalent number of
random walk histories. Each step in the random walk sequence will
contribute to the solution in the starting source state. For the
adjoint sequence, analagous to a forward stochastic transport
calculation, the source term is sampled with a set number of random
walk histories, meaning that regions of larger source will be sampled
as a starting state more regularly. In this case, each step in the
random walk sequence will contribute to the state in which it
currently resides, not the current source state. It is this set of
primary differences between the two solution methods that will dictate
solver performance with MCSA.

@@

\begin{figure}[htpb!]
  \centering
  \includegraphics[width=5in]{AdjointDirectCPUTime.png}
  \caption{\sl CPU Time (s) to converge vs. Problem Size (N for an NxN
    problem). Both the adjoint and direct solvers are used with the
    five point and nine point stencils. A siginificant CPU time
    speedup is noted with the adjoint method due to the reduced number
    of histories.}
  \label{fig:cpu_time}
\end{figure}

\begin{figure}[htbp!]
  \centering
  \includegraphics[width=5in]{AdjointDirectIterations.png}
  \caption{\sl Iterations to converge vs. Problem Size (N for an NxN
    problem). Both the adjoint and direct solvers are used with the
    five point and nine point stencils. Both methods give similar
    iteration behavior.}
  \label{fig:iterations}
\end{figure}

@@

\begin{figure}[htpb!]
  \centering
  \includegraphics[width=5in]{AdjointDirectConvergence.png}
  \caption{\sl Infinity norm of the solution residual vs. iteration
    number for a problem of fixed size. Both the adjoint and direct
    solvers are used with the five point and nine point stencils. Both
    methods give similar iteration behavior.}
  \label{fig:convergence}
\end{figure}

@@

\bibliographystyle{rnotes}
\bibliography{direct_vs_adjoint}
